Star-galaxy Classification Using Deep Convolutional Neural Networks

### 2016-06-11 paper/main.tex
add backprop citation
minor fixes and updates

\begin{abstract}
Most existing star-galaxy classifiers use the reduced summary information from catalogs,
requiring careful feature extraction and selection.
The latest advances in machine learning using deep convolutional neural networks
allow a machine to automatically learn the features directly from data,
minimizing need for input from human experts.
We present a star-galaxy classification framework that uses
a deep convolutional neural network (ConvNet) directly on the raw pixel values.
Using data from the Sloan Digital Sky Survey (SDSS) and
the Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS),
we demonstrate that ConvNet is able to produce accurate and well-calibrated
probabilistic classifications that are competitive with
conventional machine learning techniques.
Future advances in deep learning may bring more success in currently ongoing and
forthcoming photometric surveys, such as the Dark Energy Survey (DES) and
the Large Synoptic Survey Telescope (LSST), because deep neural networks require
very little feature engineering by hand.
\end{abstract}

@@

With machine learning, we can use algorithms to automatically create
accurate source catalogs with well-calibrated posterior probabilities.
Machine learning techniques have been a popular tool
in many areas of astronomy~\citep{ball2007robust,ball2008robust,
banerji2010galaxy,
carrascokind2013tpz,carrascokind2014somz,carrascokind2014exhausting,
ivezic2014statistics,
kamdar2016machine1,kamdar2016machine2}.
Artificial neural networks were first applied to the problem of star-galaxy
classification in the work of \citet{odewahn1992automated},
which has become a core part of the astronomical image processing software
\texttt{SExtractor}~\citep{bertin1996sextractor}.
Other successfully implemented examples of applying machine learning to
the star-galaxy classification problem include decision 
trees~\citep{weir1995automated,suchkov2005census,ball2006robust,sevilla2015effect}
, Support Vector Machines~\citep{Fadely2012}, and
classifier combination strategies~\citep{kim2015hybrid}.

Almost all star-galaxy classifiers published in the literature use
the reduced summary information available from astronomical catalogs.
Constructing catalogs requires careful engineering and considerable domain expertise
to transform the raw pixel values of an image to suitable features,
such as magnitudes or shape information of an object.
In a branch of machine learning called deep learning,
features are not designed by human experts;
they are learned from data by deep neural networks.
Deep learning has been applied successfully to galaxy morphological
classification in Sloan Digital Sky Survey~\citep[SDSS;][]{dieleman2015rotation}
and Cosmic Assembly Near-infrared Deep Extragalactic Legacy
Survey~\citep[CANDELS;][]{huertas2015catalog} and to photometric redshift
estimation~\citep{hoyle2015measuring},
but it has never been applied to the problem of source classification.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/sdss_color_color.pdf}
  \caption{
    The color-color space distribution of stars and galaxies in the SDSS data
    set. It is easy to see that there is significant overlap between stars
    and galaxies in the color-color space. 
  }
  \label{fig:sdss_color_color}
\end{figure}

The Sloan Digital Sky Survey~\citep[SDSS;][]{york2000sloan}
phases I--III obtained photometric data in five bands:
$u$, $g$, $r$, $i$, and $z$,
covering 14,555 square degrees, more than one-third of the entire sky.
The resulting catalog contains photometry of over 300 million stars galaxies
with a limiting magnitude of $r \approx 22$,
making the SDSS one of the largest sky surveys ever undertaken.
The SDSS also conducted an expansive spectroscopic follow-up of
more than three million stars and galaxies~\citep{eisenstein2011sdss}.
In this paper, we use a subset of the photometric and spectroscopic data
contained within the Data Release 12~\citep[DR12;][]{alam2015eleventh},
which is publicly available through the online CasJobs
server~\footnote{http://skyserver.sdss.org/casjobs/}~\citep{li2008casjobs}.

Using the CasJobs server, we randomly select $65\,000$ stars and galaxies.
We exclude some bad photometric observations as follows.
We consider only objects whose $r$ band magnitude is less than 40;
there are no warning flags in the spectroscopic measurement (\texttt{zWarning = 0});
the half-light radius in the $r$ band is less than 30 arc seconds
as measured by the exponential and de Vaucouleurs light profiles;
the error on the spectroscopic redshift measurement is less than 0.1; and
the spectroscopic redshift is less than 2.

To create training images,
we obtain the image FITS files for SDSS fields containing these objects
in five photometric bands: $u$, $g$, $r$, $i$, and $z$.
We use the astrometry information in the FITS headers
in the \texttt{Montage}~\footnote{http://montage.ipac.caltech.edu/} software
to align each image to the reference ($r$-band) image.
We then use \texttt{SExtractor} to find the pixel positions of
the $65\,000$ objects we have selected,
and to center each object in the cutout image.
We also convert the flux values to inverse hyperbolic sine magnitudes
\citep[also known as luptitudes;][]{lupton1999modified}.
Finally, in order to account for the effect of galactic dust,
galactic extinction corrections in magnitudes are applied
following \cite{schlegel1998maps}.
In the end, we have cutout images of size $48\times48$ pixels
with luptitude values in each pixel.
We note that we have experimented with increasing the pixel dimensions to
$60\times60$ and $72\times72$ pixels, but do not find noticeable
improvement in the performance of our model.

In the end, we have $17\,344$ stars and $47\,656$ galaxies available
for the training and testing processes.
The apparent magnitudes range from $10.7 < r < 23.1$,
and the galaxies in this sample have a mean redshift of $z \sim 0.36$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:sdss_color_color}.
We randomly split the objects into training, hold-out validation,
and blind test sets of size $40\,000$, $10\,000$, and $15\,000$, respectively.
We note that cross-validation is often avoided in deep learning
in favor of hold-out validation,
since cross-validation is computationally expensive.
We also note that we perform a blind test, and the test set is not used in any
way to train or calibrate the algorithms.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_color_color.pdf}
  \caption{
    The color-color space distribution of stars and galaxies in the CFHTLenS data
    set.
  }
  \label{fig:clens_color_color}
\end{figure}

@@

We obtain FITS images for each 1 square degree CFHTLenS pointing
that contains objects with spectroscopic labels.
We create cutout images of size $96\times96$ pixels by using a similar method
to that described in Section~\ref{sec:sdss}.
Finally, images are downscaled to $48\times48$ pixels
to reduce reduce computational cost.

In the end, we have $8\,545$ stars and $57\,843$ galaxies available
for the training and testing processes.
The apparent magnitudes range from $13.9 < r < 25.6$,
and the galaxies in this sample have a mean redshift of $z \sim 0.59$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:clens_color_color}.
We randomly split the objects into training, hold-out validation, and
blind test sets of size $40\,000$, $10\,000$, and $13\,278$, respectively.

@@

A neuron in the human brain receives signals from other neurons through
synaptic connections.
If the combination of these signals exceeds a certain threshold,
the neuron will fire and send a signal to other neurons.
Intelligence is believed to be the collective effect of
approximately $10^{11}$ neurons firing.
An artificial neuron in most artificial neural networks is represented
as a mathematical function that models a biological neural structure
(Figure~\ref{fig:neuron_a}).
Let $\bmath{x}=\left(x_1,x_2,\dots,x_n\right)$ be a vector of inputs to a given neuron,
$\bmath{w}=\left(w_1,w_2,\dots,w_n\right)$ be a vector of weights, and
$\bmath{b}$ be the bias.
Then, the output of the neuron is
\begin{equation}
  y = \sigma \left( \bmath{w} \cdot \bmath{x} + b \right),
  \label{eq:neuron_output}
\end{equation}
where $\sigma$ is the activation function (or \textit{non-linearity}).

@@

The goal of learning is to find a set of weights and biases such that,
given $N$ samples, the output from the network
$\bmath{y}=\left(y_1, y_2, \dots, y_N \right)$
approximates the desired output
$\hat{\bmath{y}}=\left(\hat{y}_1, \hat{y}_2, \dots, \hat{y}_N \right)$
as closely as possible for all input
$\bmath{X}=\left(\bmath{x}_1,\bmath{x}_2,\dots,\bmath{x}_N\right)$.
We can formulate this as the minimization of a loss function
$L(\bmath{y},\hat{\bmath{y}})$
over the training data.
In this work, we use \textit{cross-entropy},
\begin{equation}
  L(y_j, \hat{y}_j) = -\hat{y}_j \log_2 y_j - (1 - \hat{y}_j) \log_2(1 - y_j),
  \label{eq:cross_entropy}
\end{equation}
where $\hat{y}_j$ is the actual truth value (\eg 0 or 1) of the $j$-th data and
$y_j$ is the probability prediction made by the model.
We compute the loss function by taking the average of all cross-entropies in the sample.
Thus, the loss function becomes
\begin{equation}
  L(y_j, \hat{y}_j) = - \frac{1}{N} \sum_{j=1}^{N} \hat{y}_j  \log_2 y_j
    + (1 - \hat{y}_j)  \log_2 (1 - \hat y_j).
  \label{eq:cross_entropy_all}
\end{equation}

@@

The gradients in \ref{eq:gradient_descent} can be computed using the
\textit{backpropagation} procedure, which is nothing more than an application of
the chain rule for derivatives.
The derivative of the loss function with respect to the input of a unit can be
computed by working backwards from the gradient with respect to the output of
the unit.
At the output layer, the derivative of the loss function with respect to the
output is given by
\begin{align}
  \frac{ \partial L }{ \partial y_i } &= y_i - \hat{y}_i \\
  \frac{ \partial L }{ \partial x_i } &=
    \frac{ \partial L }{ \partial y_i }
    \frac{ \partial y_i }{ \partial x_i }.
\end{align}
At each hidden layer, the derivative of the loss function with respect to the
output of each unit is a weighted sum of the derivatives with respect to the
inputs of the subsequent layer,
\begin{align}
  \frac{ \partial L }{ \partial y_j } &=
    \sum_{i} w_{ji} \frac{ \partial L }{ \partial x_i}   \\
  \frac{ \partial L }{ \partial x_i } &=
    \frac{ \partial L }{ \partial y_i }
    \frac{ \partial y_i }{ \partial x_i }.
\end{align}
Finally, once the gradients $\partial L / \partial x_i$ have been
computed, it is straightforward to compute the gradients for the weight,
\begin{equation}
\frac{\partial L}{\partial w_{ji}} = y_j \frac{\partial L}{\partial x_i}.
\end{equation}

@@

The Convolutional neural network~\citep[ConvNet;][]{fukushima1980neocognitron,lecun1998gradient}
is a type of deep, feedforward neural network
that has recently become the dominant approach in the computer vision community.
In a typical ConvNet, the first few stages are composed of two types of layers: 
convolutional layers and pooling layers.

@@

We present the overall architecture of our ConvNet model in Table~\ref{table:hyperparamters}.
The network consists of eleven trainable layers.
The first convolutional layer filters the $5\times44\times44$ input image
(\ie $44\times44$ images in five bands $ugriz$) with 32 square filters of size $5\times5\times5$.
We have also experimented with using only three bands $gri$ (for three channels of RGB) and 
four bands $ugri$ and $griz$ (corresponding to RGBA),
and using only colors, \eg $u-g$, $g-r$, $r-i$, and/or $i-z$,
but we find that using magnitudes in all five bands $ugriz$ yields the best performance.
The leaky ReLU non-linearity is applied to the output of the first convolutional layer
(and all subsequent layers), and the second convolutional layer filters it
with 32 filters of $32\times3\times3$.
In the second convolutional layer (and all subsequent convolutional layers),
we pad the input with zeros spatially on the border
(\ie the zero-padding is 1 pixel for $3\times3$ convolutional layers)
such that the spatial resolution is preserved after convolution.
Max-pooling with filters of size $2\times2$ follows the second convolutional layer.
A stack of six additional convolutional layers, all with filters of size $3\times3$,
is followed by three fully-connected layers.
The first two fully-connected layers have 2048 channels each,
and the third performs binary classification.

@@

The choice of momentum, learning rate, and initial weights is crucial
for achieving high predictive performance and speeding up the learning
process~\citep{sutskever2013importance}.
To train our models, we use mini-batch gradient descent with a batch size of 128
and Nesterov momentum~\citep{bengio2013advances} of 0.9.
We initialize the learning rate $\eta$ at 0.003 for all layers and
decrease it linearly with the number of epochs 
from 0.003 to 0.0001 over 750 epochs.
We also initialize the weights in each layer with random orthogonal initial 
conditions~\citep{saxe2013exact}.
We use a slightly positive value $b=0.1$ for all biases.
We find initializing biases to a small constant value helps eliminate dead neurons
by ensuring that all ReLU neurons fire in the beginning.

We note that our network takes about 40 hours to train the networks on a NVIDIA Tesla K40 GPU.

@@

\item{Reflection:
We flip each image horizontally with a probability of 0.5 to exploit mirror symmetry. }
\item{Translation:
We also have translational symmetry in the images.
Given an image of size $48\times48$ pixels, we extract a random contiguous crop
of size $44\times44$.
Each cropping is equivalent to randomly shifting a $44\times44$ image by up to 4 pixels
vertically and horizontally. }
\item{Gaussian noise:
We introduce random Gaussian noise to each pixel values
by using a similar method to \cite{krizhevsky2012imagenet}. }
\end{itemize}
These data augmentation schemes add almost no computational cost,
as they are performed on the CPU while the GPU is training the ConvNets on images.

@@

In this section, we first describe the performance metrics we use to evaluate
the models. We then present the classification performance of our ConvNet model
on CFHTLenS and SDSS data sets, and compare it with the performance of TPC.

@@

\begin{table}
  \caption{The definition of the classification performance metrics.}
  \centering
  \begin{tabular}{c l}
  \hline
  Metric & Meaning \\
  \hline
  AUC & Area under the Receiver Operating Curve \\
  MSE & Mean squared error \\
  $c_g$ & Galaxy completeness \\
  $p_g$ & Galaxy purity \\
  $c_s$ & Star completeness \\
  $p_s$ & Star purity \\
  $p_g(c_g=x)$ & Galaxy purity at $x$ galaxy completeness \\
  $p_s(c_s=x)$ & Star purity at $x$ star completeness \\
  $CAL$ & Calibration error with overlapping binning \\
  $\mid\Delta N_g\mid / N_g$ & Absolute error in number of galaxies\\
  \hline
  \end{tabular}
  \label{table:metrics}
\end{table}

@@

We also use the mean squared error
(MSE; also known as Brier score~\citep{brier1950verification})
as a performance metric. We define MSE as
\begin{equation} \label{eq:mse}
  \rmn{MSE} = \frac{1}{N} \sum^{N}_{j=1}
  \left( y_j - \hat{y}_j \right)^2,
\end{equation}
The MSE can be considered as both a score function that quantifies
how well a set of probabilistic predictions are calibrated,
or a loss function.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_mag.pdf}
  \caption{Galaxy completeness and star purity values as functions
           of the $i$-band magnitude
           as estimated by the kernel density estimation (KDE) method
           for the differential counts in the CFHTLenS data set.
           The top panel shows the histogram with a bin size of 0.1 mag
           and the KDE for objects in the test set.
           The second panel shows the fraction of stars estimated by KDE
           as a function of magnitude.
           The bottom two panels compare the galaxy purity and star completeness
           values for ConvNet (red solid line),
           $\rmn{TPC}_{\rmn{morph}}$ (blue dashed line),
           and $\rmn{TPC}_{\rmn{phot}}$ (yellow dashed line)
           as functions of magnitude.
           The $1 \sigma$ confidence bands are estimated by
           bootstrap sampling.}
  \label{fig:clens_mag}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_integrated.pdf}
  \caption{
    Galaxy purity and star completeness as functions of the $i$-band
    magnitude for the integrated counts in the CFHTLenS data set.
    The upper panel compares the galaxy
    purity values for ConvNet (red solid line), $\rmn{TPC}_{\rmn{morph}}$
    (blue dashed line), and $\rmn{TPC}_{\rmn{phot}}$ (yellow dashed line).
    The lower panel compares the star completeness.
    The $1 \sigma$ error bars are computed following the method
    of \citet{paterno2004calculating} to avoid the unphysical
    errors of binomial or Poisson statistics.
    }
  \label{fig:clens_integrated}
\end{figure}

@@

\begin{figure*}
  \centering
  \begin{minipage}[c]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/clens_calibration_conv.pdf}
    \caption{
      Calibration curve for ConvNet as applied to the CFHTLenS data set.
      $P_{\rmn{gal}}$ is the fraction of objects that are galaxies.
      $P_{\rmn{conv}}$ is the probabilistic output generated by ConvNet.
      The dashed line displays the relationship
      $P_{\rmn{gal}} = P_{\rmn{conv}}$.
      The $1 \sigma$ error bars are computed following the method
      of \citet{paterno2004calculating}.
      }
    \label{fig:clens_calibration_conv}
  \end{minipage}
  \hfill
  \begin{minipage}[c]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/clens_calibration_tpc.pdf}
    \caption{Similar to Figure~\ref{fig:clens_calibration_conv} but for
      $\rmn{TPC}_{\rmn{morph}}$.
      $P_{\rmn{tpc}}$ is the probabilistic output generated by ConvNet.}
    \label{fig:clens_calibration_tpc}
  \end{minipage}
\end{figure*}

@@

As descrbied in Section~\ref{sec:convnet},
we train our ConvNet model by monitoring its performance on the validatoin set.
Once we have finished training our model,
we evaluate its performance on the blind test set.
We also use the same training and validation sets to train 
$\rmn{TPC}_{\rmn{morph}}$ and $\rmn{TPC}_{\rmn{phot}}$, and
perform classifications on the same test set to compare their performance
with that of ConvNet.

>>>>

\begin{abstract}
Most existing star-galaxy classifiers use the reduced summary information from catalogs,
requiring careful feature extraction and selection.
The latest advances in machine learning using deep convolutional neural networks
allow a machine to automatically learn the features directly from data,
minimizing need for input from human experts.
We present a star-galaxy classification framework that uses
deep convolutional neural networks (ConvNets) directly on the raw pixel values.
Using data from the Sloan Digital Sky Survey (SDSS) and
the Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS),
we demonstrate that ConvNets are able to produce accurate and well-calibrated
probabilistic classifications that are competitive with
conventional machine learning techniques.
Future advances in deep learning may bring more success in currently ongoing and
forthcoming photometric surveys, such as the Dark Energy Survey (DES) and
the Large Synoptic Survey Telescope (LSST), because deep neural networks require
very little feature engineering by hand.
\end{abstract}

@@

With machine learning, we can use algorithms to automatically create
accurate source catalogs with well-calibrated posterior probabilities.
Machine learning techniques have been a popular tool
in many areas of astronomy~\citep{ball2007robust,ball2008robust,
banerji2010galaxy,
carrascokind2013tpz,carrascokind2014somz,carrascokind2014exhausting,
ivezic2014statistics,
kamdar2016machine1,kamdar2016machine2}, and
artificial neural networks were first applied to the problem of star-galaxy
classification in the work of \citet{odewahn1992automated},
which has become a core part of the astronomical image processing software
\texttt{SExtractor}~\citep{bertin1996sextractor}.
Other successfully implemented examples of applying machine learning to
the star-galaxy classification problem include decision 
trees~\citep{weir1995automated,suchkov2005census,ball2006robust,sevilla2015effect}
, Support Vector Machines~\citep{Fadely2012}, and
classifier combination strategies~\citep{kim2015hybrid}.

Almost all star-galaxy classifiers published in the literature use
the reduced summary information available from astronomical catalogs.
Constructing catalogs requires careful engineering and considerable domain expertise
to transform the raw pixel values of an image to suitable features,
such as magnitudes or shape information of an object.
In a branch of machine learning called \textit{deep learning},
features are not designed by human experts;
they are learned directly from data by deep neural networks.
Deep learning has been applied successfully to galaxy morphological
classification in Sloan Digital Sky Survey~\citep[SDSS;][]{dieleman2015rotation}
and Cosmic Assembly Near-infrared Deep Extragalactic Legacy
Survey~\citep[CANDELS;][]{huertas2015catalog} and to photometric redshift
estimation~\citep{hoyle2015measuring},
but it has never been applied to the problem of source classification.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/sdss_color_color.pdf}
  \caption{
    Hexbin plots showing the color-color space distribution of stars and
    galaxies in the SDSS data set.
    It is easy to see that there is significant overlap between stars
    and galaxies in the color-color space. 
  }
  \label{fig:sdss_color_color}
\end{figure}

The Sloan Digital Sky Survey~\citep[SDSS;][]{york2000sloan}
phases I--III obtained photometric data in five bands,
$u$, $g$, $r$, $i$, and $z$,
covering 14,555 square degrees, more than one-third of the entire sky.
The resulting catalog contains photometry of over 300 million stars galaxies
with a limiting magnitude of $r \approx 22$,
making the SDSS one of the largest sky surveys ever undertaken.
The SDSS also conducted an expansive spectroscopic follow-up of
more than three million stars and galaxies~\citep{eisenstein2011sdss}.
In this paper, we use a subset of the photometric and spectroscopic data
contained within the Data Release 12~\citep[DR12;][]{alam2015eleventh},
which is publicly available through the online CasJobs
server\footnote{http://skyserver.sdss.org/casjobs/}~\citep{li2008casjobs}.

Using the CasJobs server, we randomly select $65\,000$ stars and galaxies.
We exclude some bad photometric observations as follows.
We consider only objects whose $r$-band magnitude is less than 40;
there are no warning flags in the spectroscopic measurement (\texttt{zWarning = 0});
the half-light radius in the $r$ band is less than 30 arc seconds
as measured by the exponential and de Vaucouleurs light profiles;
the error on the spectroscopic redshift measurement is less than 0.1; and
the spectroscopic redshift is less than 2.

To create training images,
we obtain the image FITS files for SDSS fields containing these objects
in five photometric bands: $u$, $g$, $r$, $i$, and $z$.
We use the astrometry information in the FITS headers
in the \texttt{Montage}\footnote{http://montage.ipac.caltech.edu/} software
to align each image to the reference ($r$-band) image.
We then use \texttt{SExtractor} to find the pixel positions of
the $65\,000$ objects we have selected,
and to center each object in the cutout image.
We also convert the flux values to inverse hyperbolic sine magnitudes
\citep[also known as luptitudes;][]{lupton1999modified}.
Finally, in order to account for the effect of galactic dust,
galactic extinction corrections in magnitudes are applied
following \cite{schlegel1998maps}.
In the end, we have cutout images of size $48\times48$ pixels
with luptitude values in each pixel.
We note that we have experimented with increasing the pixel dimensions to
$60\times60$ and $72\times72$ pixels, but do not find noticeable
improvement in the performance of our model.

In the end, we have $17\,344$ stars and $47\,656$ galaxies available
for the training and testing processes.
The apparent magnitudes range from $10.7 < r < 23.1$,
and the galaxies in this sample have a mean redshift of $z \sim 0.36$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:sdss_color_color}.
We randomly split the objects into training, held-out validation,
and blind test sets of size $40\,000$, $10\,000$, and $15\,000$, respectively.
We note that cross-validation is often avoided in deep learning
in favor of hold-out validation,
since cross-validation is computationally expensive.
We also note that we perform a blind test, and the test set is not used in any
way to train or calibrate the algorithms.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_color_color.pdf}
  \caption{
    Hexbin plots showing the color-color space distribution of stars and
    galaxies in the CFHTLenS data set.
  }
  \label{fig:clens_color_color}
\end{figure}

@@

We obtain FITS images for each 1 square degree CFHTLenS pointing
that contains objects with spectroscopic labels.
We create cutout images of size $96\times96$ pixels by using a similar method
to that described in Section~\ref{sec:sdss}.
Finally, images are downscaled to $48\times48$ pixels
to reduce computational cost.

In the end, we have $8\,545$ stars and $57\,843$ galaxies available
for the training and testing processes.
The apparent magnitudes range from $13.9 < r < 25.6$,
and the galaxies in this sample have a mean redshift of $z \sim 0.59$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:clens_color_color}.
We randomly split the objects into training, held-out validation, and
blind test sets of size $40\,000$, $10\,000$, and $13\,278$, respectively.

@@

A neuron in the human brain receives signals from other neurons through
synaptic connections.
If the combination of these signals exceeds a certain threshold,
the neuron will fire and send a signal to other neurons.
Intelligence is believed to be the collective effect of
approximately $10^{11}$ neurons firing.
An artificial neuron in most artificial neural networks is represented
as a mathematical function that models a biological neural structure
(Figure~\ref{fig:neuron_a}).
Let $\bmath{x}=\left(x_1,x_2,\dots,x_n\right)$ be a vector of inputs to a given neuron,
$\bmath{w}=\left(w_1,w_2,\dots,w_n\right)$ be a vector of weights, and
$b$ be the bias.
Then, the output of the neuron is
\begin{equation}
  y = \sigma \left( \bmath{w} \cdot \bmath{x} + b \right),
  \label{eq:neuron_output}
\end{equation}
where $\sigma$ is the activation function (or \textit{non-linearity}).

@@

The goal of learning is to find a set of weights and biases such that,
given $N$ samples, the output from the network
$\bmath{y}=\left(y_1, y_2, \dots, y_N \right)$
approximates the desired output
$\hat{\bmath{y}}=\left(\hat{y}_1, \hat{y}_2, \dots, \hat{y}_N \right)$
as closely as possible for all input
$\bmath{X}=\left(\bmath{x}_1,\bmath{x}_2,\dots,\bmath{x}_N\right)$.
We can formulate this as the minimization of a loss function
$L(\bmath{y},\hat{\bmath{y}})$
over the training data.
In this work, we use \textit{cross-entropy},
\begin{equation}
  L(y_j, \hat{y}_j) = -\hat{y}_j \log_2 y_j - (1 - \hat{y}_j) \log_2(1 - y_j),
  \label{eq:cross_entropy}
\end{equation}
where $\hat{y}_j$ is the actual truth value (\eg 0 or 1) of the $j$-th data, and
$y_j$ is the probability prediction made by the model.
We compute the loss function by taking the average of all cross-entropies in the sample.
Thus, the loss function becomes
\begin{equation}
  L(y_j, \hat{y}_j) = - \frac{1}{N} \sum_{j=1}^{N} \hat{y}_j  \log_2 y_j
    + (1 - \hat{y}_j)  \log_2 (1 - \hat y_j).
  \label{eq:cross_entropy_all}
\end{equation}

@@

The gradients in \ref{eq:gradient_descent} can be computed using the
\textit{backpropagation} procedure~\citep{rumelhart1988learning},
which is nothing more than an application of the chain rule for derivatives.
The derivative of the loss function with respect to the input of a unit can be
computed by working backwards from the gradient with respect to the output of
the unit.
At the output layer, the derivative of the loss function with respect to the
output is given by
\begin{align}
  \frac{ \partial L }{ \partial y_i } &= y_i - \hat{y}_i \\
  \frac{ \partial L }{ \partial x_i } &=
    \frac{ \partial L }{ \partial y_i }
    \frac{ \partial y_i }{ \partial x_i }.
\end{align}
At each hidden layer, the derivative of the loss function with respect to the
output of each unit is a weighted sum of the derivatives with respect to the
inputs of the subsequent layer,
\begin{align}
  \frac{ \partial L }{ \partial y_j } &=
    \sum_{i} w_{ji} \frac{ \partial L }{ \partial x_i}   \\
  \frac{ \partial L }{ \partial x_j } &=
    \frac{ \partial L }{ \partial y_j }
    \frac{ \partial y_j }{ \partial x_j }.
\end{align}
Finally, once the gradients $\partial L / \partial x_j$ have been
computed, it is straightforward to compute the gradients for the weight,
\begin{equation}
\frac{\partial L}{\partial w_{jk}} = y_k \frac{\partial L}{\partial x_j}.
\end{equation}

@@

The convolutional neural network~\citep[ConvNet;][]{fukushima1980neocognitron,lecun1998gradient}
is a type of deep, feedforward neural network
that has recently become the dominant approach in the computer vision community.
In a typical ConvNet, the first few stages are composed of two types of layers: 
convolutional layers and pooling layers.

@@

We present the overall architecture of our ConvNet model in Table~\ref{table:hyperparamters}.
The network consists of eleven trainable layers.
The first convolutional layer filters the $5\times44\times44$ input image
(\ie $44\times44$ images in five bands $ugriz$) with 32 square filters of size $5\times5\times5$.
We have also experimented with using only three bands $gri$ (for three channels of RGB) and 
four bands $ugri$ and $griz$ (corresponding to RGBA),
and using only colors, \eg $u-g$, $g-r$, $r-i$, and/or $i-z$,
but we find that using magnitudes in all five bands $ugriz$ yields the best performance.

The leaky ReLU non-linearity is applied to the output of the first convolutional layer
(and all subsequent layers), and the second convolutional layer filters it
with 32 filters of $32\times3\times3$.
In the second convolutional layer (and all subsequent convolutional layers),
we pad the input with zeros spatially on the border
(\ie the zero-padding is 1 pixel for $3\times3$ convolutional layers)
such that the spatial resolution is preserved after convolution.
Max-pooling with filters of size $2\times2$ follows the second convolutional layer.
A stack of six additional convolutional layers, all with filters of size $3\times3$,
is followed by three fully-connected layers.
The first two fully-connected layers have 2048 channels each,
and the third performs binary classification.

@@

The choice of momentum, learning rate, and initial weights is crucial
for achieving high predictive performance and speeding up the learning
process~\citep{sutskever2013importance}.
To train our models, we use mini-batch gradient descent with a batch size of 128
and Nesterov momentum~\citep{bengio2013advances} of 0.9.
We initialize the learning rate $\eta$ at 0.003 for all layers and
decrease it linearly with the number of epochs 
from 0.003 to 0.0001 over 750 epochs.
We also initialize the weights in each layer with random orthogonal initial 
conditions~\citep{saxe2013exact}.
We use slightly positive values ($b=0.01$ or $0.1$) for all biases.
We find initializing biases to a small constant value helps eliminate dead neurons
by ensuring that all ReLU neurons fire in the beginning.

We note that training our network takes about 40 hours on a NVIDIA Tesla K40 GPU.

@@

\item{Reflection:
We flip each image horizontally with a probability of 0.5 to exploit mirror symmetry. }
\item{Translation:
We also have translational symmetry in the images.
Given an image of size $48\times48$ pixels, we extract a random contiguous crop
of size $44\times44$.
Each cropping is equivalent to randomly shifting a $44\times44$ image by up to 4 pixels
vertically and/or horizontally. }
\item{Gaussian noise:
We introduce random Gaussian noise to each pixel values
by using a similar method to \cite{krizhevsky2012imagenet}. }
\end{itemize}
These data augmentation schemes add almost no computational cost,
as they are performed on the CPU while the GPU is training the ConvNets on images.

@@

In this section, we first describe the performance metrics that were used for
evaluating the models.
We then present the classification performance of our ConvNet model
on the CFHTLenS and SDSS data sets, and compare it with the performance of TPC.

@@

\begin{table}
  \caption{The definition of the classification performance metrics.}
  \centering
  \begin{tabular}{c l}
  \hline
  Metric & Meaning \\
  \hline
  AUC & Area under the Receiver Operating Curve \\
  MSE & Mean squared error \\
  $c_g$ & Galaxy completeness \\
  $p_g$ & Galaxy purity \\
  $c_s$ & Star completeness \\
  $p_s$ & Star purity \\
  $p_g(c_g=x)$ & Galaxy purity at $x$ galaxy completeness \\
  $c_s(p_s=x)$ & Star completeness at $x$ star purity \\
  $CAL$ & Calibration error with overlapping binning \\
  $\mid\Delta N_g\mid / N_g$ & Absolute error in number of galaxies\\
  \hline
  \end{tabular}
  \label{table:metrics}
\end{table}

@@

We also use the mean squared error
(MSE; also known as Brier score~\citep{brier1950verification})
as a performance metric. We define MSE as
\begin{equation} \label{eq:mse}
  \rmn{MSE} = \frac{1}{N} \sum^{N}_{j=1}
  \left( y_j - \hat{y}_j \right)^2,
\end{equation}
The MSE can be considered as both a score function that quantifies
how well a set of probabilistic predictions is calibrated,
or a loss function.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_mag.pdf}
  \caption{Galaxy purity and star completeness values as functions
           of the $i$-band magnitude
           as estimated by the kernel density estimation (KDE) method
           for the differential counts in the CFHTLenS data set.
           The top panel shows the histogram with a bin size of 0.1 mag
           and the KDE for objects in the test set.
           The second panel shows the fraction of stars estimated by KDE
           as a function of magnitude.
           The bottom two panels compare the galaxy purity and star completeness
           values for ConvNet (red solid line),
           $\rmn{TPC}_{\rmn{morph}}$ (blue dashed line),
           and $\rmn{TPC}_{\rmn{phot}}$ (yellow dashed line)
           as functions of magnitude.
           The $1 \sigma$ confidence bands are estimated by
           bootstrap sampling.}
  \label{fig:clens_mag}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_integrated.pdf}
  \caption{
    Galaxy purity and star completeness as functions of the $i$-band
    magnitude for the integrated counts in the CFHTLenS data set.
    The upper panel compares the galaxy
    purity values for ConvNet (red solid line), $\rmn{TPC}_{\rmn{morph}}$
    (blue dashed line), and $\rmn{TPC}_{\rmn{phot}}$ (yellow dashed line).
    The lower panel compares the star completeness values.
    The $1 \sigma$ error bars are computed following the method
    of \citet{paterno2004calculating} to avoid the unphysical
    errors of binomial or Poisson statistics.
    }
  \label{fig:clens_integrated}
\end{figure}

@@

\begin{figure*}
  \centering
  \begin{minipage}[c]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/clens_calibration_conv.pdf}
    \caption{
      Calibration curve for ConvNet as applied to the CFHTLenS data set.
      $P_{\rmn{gal}}$ is the fraction of objects that are galaxies.
      $P_{\rmn{conv}}$ is the probabilistic output generated by ConvNet.
      The dashed line displays the relationship
      $P_{\rmn{gal}} = P_{\rmn{conv}}$.
      The $1 \sigma$ error bars are computed following the method
      of \citet{paterno2004calculating}.
      }
    \label{fig:clens_calibration_conv}
  \end{minipage}
  \hfill
  \begin{minipage}[c]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figures/clens_calibration_tpc.pdf}
    \caption{Similar to Figure~\ref{fig:clens_calibration_conv} but for
      $\rmn{TPC}_{\rmn{morph}}$.
      $P_{\rmn{tpc}}$ is the probabilistic output generated by $\rmn{TPC}_{\rmn{morph}}$.}
    \label{fig:clens_calibration_tpc}
  \end{minipage}
\end{figure*}

@@

As described in Section~\ref{sec:convnet},
we train our ConvNet model by monitoring its performance on the validation set.
Once we have finished training the model,
we evaluate its performance on the blind test set.
We also use the same training and validation sets to train 
$\rmn{TPC}_{\rmn{morph}}$ and $\rmn{TPC}_{\rmn{phot}}$, and
perform classifications on the same test set to compare their performance
with that of ConvNet.
